---
title: "Stat 184: Final Project"
author: "Ally Racho"
due date: "12/10/2020"
output: html_document
---

## Front Matter

### Necessary Libraries:
```{r, message=FALSE}
library(mosaic)
library(tidyverse)
library(readxl)
```

### Imported Datasets:
```{r, message=FALSE, warning=FALSE}
read.csv("C:\\Users\\allyracho\\Desktop\\NYPD_Complaint_Data_Historic.csv") # Import primary source
read_excel("C:\\Users\\allyracho\\Desktop\\MeanZIP-3.xlsx") # Import secondary source
```
Since the original dataset has over 1 million cases and is too large of a file, I will be sampling my data to a smaller set of 250000 cases. 
```{r}
SampleData <-
  sample(NYPD_Complaint_Data_Historic, 250000)
```

## Becoming Acquainted with data:

1. NYPD data:
```{r}
names(SampleData)     ## Names of variables
nrow(SampleData)       ## Number of Cases
str(SampleData)       ## Structure of object, information of class
```

2. MeanZip Data:
```{r}
names(MeanZIP_3)     ## Names of variables
nrow(MeanZIP_3)       ## Number of Cases
str(MeanZIP_3)       ## Structure of object, information of class
```

## Research Question

Based on the NYC borough and average income per borough, does there seem to be a trend in a certain type of crime committed there?

## Wrangling and Plots

Most popular crime based on borough (NYPD) -- figure out most popular crime in boroughs
```{r}
PopularBoroughCrime <-
  SampleData %>%
  group_by(BORO_NM, OFNS_DESC) %>%
  summarise(total = sum(n())) %>%
  filter(total == max(total))

PopularBoroughCrime %>%
  ggplot(aes(x = BORO_NM, y = total, fill = OFNS_DESC))+
  geom_bar(stat= 'identity', position = 'dodge', width = .9)
```

Based on zip code of borough, what is the mean family income? (MeanZip_3) -- find mean income per borough 
```{r}
MedianIncome <- 
  MeanZIP_3 %>%
  filter(Zip == 10453 | Zip == 11212 | Zip == 10004 | Zip == 11361 | Zip == 10314) %>%
  ggplot(aes(x = Zip, y =Mean))+
  geom_point()
```

Most popular crime over time per borough (NYPD)
```{r}
year <- "[0-9]{4}"


PopularBoroughCrimePerYear <-
  SampleData %>%
  group_by(BORO_NM, OFNS_DESC, year)%>%
  summarise(total = sum(n())) %>%
  filter(total == max(total))

PopularBoroughCrimePerYear %>%
  ggplot(aes(x = CMPLNT_FR_DT, y = total, color = OFNS_DESC)) +
  geom_point() +
  geom_line()
PopularBoroughCrimePerYear
```

Join Popular crime over time and mena income



