---
title: "NYPD Complaint Data: STAT 184 Final Project" 
author: "Ally Racho"
output: html_notebook
---

Necessary Libraries:
```{r}
library(mosaic)
library(tidyverse)
```

Since the original dataset has over 1 million cases and is too large of a file, I will be sampling my data to a smaller set of 250000 cases. 
```{r}
SampleData <-
  sample(NYPD_Complaint_Data_Historic, 250000)
```


### Research Question

Based on the NYC borough and average income per borough, does there seem to be a trend in a certain type of crime committed there?


### Becoming Acquainted with data:

-  Where did you find them? 
  -- www.data.cityofnewyork.us,www.psc.isr.umich.edu/dis/census/Features/tract2zip/

-  Who collected/maintains them? 
  -- The dataset for NYPD_Complaint_Data_Historic is provided by the NYPD and maintained/updated by NYC Open Data. The data for MeanZIP is provided and maintained by University of Michigan Population Studies Center. 

-  When & Why were they originally collected?
  -- The NYPD data was originally collected to engage New Yorkers in information that is produced and collected by the City government. The collection started in 2006, but was made public in 2016 and is updated frequently. The MeanZip was created to explain how to create a zipcode dataset. The data is based on information from 2006-2010. I was unable to see when the data was originally released by the website. 

-  What does a case represent in each data source, and how many total cases are available?
  -- A case for NYPD represents a crime that occurred on a certain day, time, location. In the original dataset there are 6.5 million cases, but I created a sample subset which contains 250000. A case for MeanZip is based on zipcode, and the avg/median income of such. There are 32634 cases for this dataset.

-  What are some of the variables that you plan to use
  -- For the NYPD dataset, I plan to use the BORO_NM (borough crime occurred in), OFNS_DESC (crime description), CMPLNT_FR_DT (exact time of occurence), and LAW_CAT_CD (level of offense, ie. felony, misdemeanor). For MeanZIP, I plan to use Zip (Zip code), Mean (Avg. Income), Pop (Population). 

```{r}
names(SampleData)     ## Names of variables
nrow(SampleData)       ## Number of Cases
str(SampleData)       ## Structure of object, information of class
```

```{r}
names(MeanZIP_3)     ## Names of variables
nrow(MeanZIP_3)       ## Number of Cases
str(MeanZIP_3)       ## Structure of object, information of class
```

## Exploring intuition related to the research question

### Create some informative plots and summary statistics 

1. Level of crime faceted among the boroughs (NYPD)
```{r}
FelonyMisdCrimeBasedonBorough <-
  SampleData %>%
  group_by(BORO_NM,LAW_CAT_CD) %>%
  summarize(total = sum(n()))

FelonyMisdCrimeBasedonBorough %>%
  ggplot(aes(x = LAW_CAT_CD, y = total, fill = LAW_CAT_CD))+
  geom_bar(stat= 'identity', position = 'dodge', width = .9)+
  facet_wrap(~ BORO_NM, ncol=3)

```

2. Most popular crime based on borough (NYPD)
```{r}
PopularBoroughCrime <-
  SampleData %>%
  group_by(BORO_NM, OFNS_DESC) %>%
  summarise(total = sum(n())) %>%
  filter(total == max(total))

PopularBoroughCrime %>%
  ggplot(aes(x = BORO_NM, y = total, fill = OFNS_DESC))+
  geom_bar(stat= 'identity', position = 'dodge', width = .9)
```

3. Based on zip code of borough, what is the median family income (MeanZip_3)
```{r}
MedianIncome <- 
  MeanZIP_3 %>%
  filter(Zip == 10453 | Zip == 11212 | Zip == 10004 | Zip == 11361 | Zip == 10314) %>%
  ggplot(aes(x = Zip, y =Mean))+
  geom_point()
MedianIncome
```



### Describe preliminary observations and intuition about the research question
* Plot 1: You can see from the distribution, the most popular level of crime is misdemeanor, with felony second. Brooklyn appears to have the highest amount of crime, and Staten Island the lowest.
* Plot 2: The crime that happens the most often is petit larceny, with the maximum total amount of times approximately ~20000 in Manhattan. Harrassment 2 is the leading crime in staten island with approximately ~2500 occurences. 
* Plot 3: You can see based on Zipcode, 10004 (Manhattan) has the highest mean income at approximately ~160000 and the lowest being 10453 (Bronx) and 11212 (Queens).  


### Two Data Sources

- Primary data: NYPD_Complaint_Data_Historic --- SampleData

- Other data sources: MeanZIP-3






